---
title: "Format input files"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Make metadata file

Downloaded data from:

https://datagateway.nrcs.usda.gov/

A way to group points by shapes they are in:

https://gis.stackexchange.com/questions/133625/checking-if-points-fall-within-polygon-shapefile

```{r}
library(rgeos)
library(sp)
library(sf)
library(rgdal)
library(tidyverse)
library(ggplot2)
library(viridis)
metadata <- read_csv('data/curry_county_population_data_site_host_nextstrain.csv')
```

### Plot data over watershed boundries

https://stackoverflow.com/questions/18174703/spatialpolygondataframe-plotting-using-ggplot

```{r}
polys <- st_read('data/hydrologic_units_WBDHU12_mbr_3971222_03/hydrologic_units/wbdhu12_a_mbr.shp')
# polys <- st_read('data/hydrologic_units_WBDHU10_mbr_3971222_02/hydrologic_units/wbdhu10_a_mbr.shp')
ggplot(polys) +
  geom_sf() + 
  geom_point(data = metadata, mapping = aes(x = lon, y = lat)) + 
  xlim(-124.5, -124.2) +
  ylim(42.05, 42.4)
```

Even the finest resolution watershed info available is not very detailed compared to the sample extent.


### find which watersheds each sample is in

```{r}
watersheds <- readOGR('data/hydrologic_units_WBDHU12_mbr_3971222_03/hydrologic_units/wbdhu12_a_mbr.shp')
metadata <- read_csv('data/curry_county_population_data_site_host_nextstrain.csv')

# Assignment modified according
meta_subset <- filter(metadata, !is.na(lon))
coordinates(meta_subset) <- ~ lon + lat
# Set the projection of the SpatialPointsDataFrame using the projection of the shapefile
proj4string(meta_subset) <- proj4string(watersheds)
meta_subset_placements <- as_tibble(over(meta_subset, watersheds))
# Combine with metadata
meta_subset$watershed <- meta_subset_placements$name
metadata <- left_join(metadata, as_tibble(meta_subset)[, c("strain", "watershed")], by = "strain")
```

find centroid coordinates of each watershed

```{r}
centroid_data <- as_tibble(gCentroid(watersheds, byid = TRUE))
colnames(centroid_data) <- c("watershed_lon", "watershed_lat")
centroid_data <- bind_cols(watershed = watersheds@data$name, centroid_data)
metadata <- left_join(metadata, centroid_data, by = "watershed")
```


add a finer resolution classification based on grids

```{r}
range(metadata$lat, na.rm = TRUE)
range(metadata$lon, na.rm = TRUE)
diff(range(metadata$lat, na.rm = TRUE))
diff(range(metadata$lon, na.rm = TRUE))
metadata$rounded_lat <- round(metadata$lat, digits = 2)
metadata$rounded_lon <- round(metadata$lon, digits = 2)
metadata$grid <- paste0(metadata$rounded_lat, ', ', metadata$rounded_lon)
metadata$grid[metadata$grid == "NA, NA"] <- NA
```

fix date format

```{r}
metadata$date <- paste0(metadata$year, '-01-01')
```

save metadata file

```{r}
write_tsv(metadata, file = file.path('data', 'metadata_modified.tsv'))
```


## Make coordinate file

```{r}
watershed_data <- metadata %>%
  filter(!is.na(watershed)) %>%
  mutate(group = "watershed") %>%
  select(group, name = watershed, lat = watershed_lat, lon = watershed_lon) %>%
  distinct()
grid_data <- metadata %>%
  filter(!is.na(grid)) %>%
  mutate(group = "grid") %>%
  select(group, name = grid, lat = rounded_lat,  lon = rounded_lon) %>%
  distinct()
coord_data <- bind_rows(watershed_data, grid_data)
write_tsv(coord_data, file = file.path("config", "lat_longs.tsv"), col_names = FALSE)
```


## Make dropped strains file

These will be filtered out of the analysis.
First, I will remove the reference strain, since that gets assigned a date in 1980 for some reason and that is confusing:

```{r}
dropped_ids <- c('MT_DQ832718.1')
```

And all IDs associated with a missing location:

```{r}
dropped_ids <- c(dropped_ids, metadata$strain[is.na(metadata$lon)])
```

And write the file, one ID per line:

```{r}
write_lines(dropped_ids, file = file.path("config", "dropped_strains.txt"))
```


## Make color file

```{r}
color_data <- coord_data %>%
  select(group, name) %>%
  mutate(color = viridis(length(group)))
color_data$color <- substr(color_data$color, start = 1, stop = 7)
write_tsv(color_data, file = file.path("config", "colors.tsv"), col_names = FALSE)
```

